\documentclass{homework}

\title{PS5}
\author{Luke Denton}

\begin{document}

\maketitle

\exercise
For my first piece of data, I pulled the top 25 long jumps ever recorded in the world by men. Although this data doesn't directly pertain to any of my potential research topics for this semester, I do find it interesting. The record, having stood for 30 years, is a remarkable feat. I like to follow track and field at the Olympics and some national championships, and these ridiculous performances all seem to largely exclude the 2010s decade. I think this is an interesting phenomenon, and I am curious as to the explanation for this. To parse the data, I used the inspector within google chrome to hover over the table in wikipedia and copy its html location. From there, I just used the html table function from rvest to get my data formatted into a data frame. 
\exercise*
For my data, I decided to look at unemployment data from the fred. I have understood that unemployment during the pandemic was at a nearly record high, however I was unsure as to how it compared to the unemployment rates of each month over my lifetime. Although there was an economic recession in 2008-2009, the unemployment rate returned to 5 percent by Sept 2015, roughly 7 years after the recession. I am curious about the recovery from the current recession, and what specifically will be the driver behind the recovery. I think the aggressive public policy will be interesting to consider as time goes on with changes in the unemployment rate and also the poverty rate. I used the fredr package that we went over in class to parse this table.

\end{document}
