\documentclass{homework}

\title{I was planning to go to Peru this week, things went awry. Going to see Godzilla vs Kong instead.}
\author{PS7 - Luke Denton}

\begin{document}

\maketitle

\exercise
logwage is missing at 25 percent, and I think that this variable is missing not at random.
\begin{table}[h]
\centering
\begin{tabular}[t]{lrrrrrrr}
\toprule
  & Unique (\#) & Missing (\%) & Mean & SD & Min & Median & Max\\
\midrule
logwage & 670 & 25 & 1.6 & 0.4 & 0.0 & 1.7 & 2.3\\
hgc & 16 & 0 & 13.1 & 2.5 & 0 & 12.0 & 18\\
tenure & 259 & 0 & 6.0 & 5.5 & 0.0 & 3.8 & 25.9\\
age & 13 & 0 & 39.2 & 3.1 & 34 & 39.0 & 46\\
\bottomrule
\end{tabular}
\end{table}

\exercise
The iterated deletion of the na's for model 1 actually got closer to the true value of Beta 1 of hgc than Model 2 or 3, which I think is very interesting. Mean imputation was by far the worst method for calculating our Beta 1 value. Using predicted values got us close to the value we obtained for both model 1 and 4. The estimates in Model 3 and 4 were similar, although our multiple imputation model did a better job of predicting the true value of Beta 1. 

\begin{table}[h]
\centering
\begin{tabular}[t]{lcccc}
\toprule
  & Model 1 & Model 2 & Model 3 & Model 4\\
\midrule
(Intercept) & 0.534 & 0.708 & 0.563 & 0.495\\
 & (0.146) & (0.116) & (0.112) & (0.137)\\
hgc & 0.062 & 0.050 & 0.059 & 0.064\\
 & (0.005) & (0.004) & (0.004) & \vphantom{1} (0.005)\\
collegenot college grad & 0.145 & 0.169 & 0.178 & 0.145\\
 & (0.034) & (0.026) & (0.025) & (0.032)\\
tenure & 0.050 & 0.038 & 0.047 & 0.050\\
 & (0.005) & (0.004) & (0.004) & (0.005)\\
tenuresq & -0.002 & -0.001 & -0.002 & -0.002\\
 & (0.000) & (0.000) & (0.000) & (0.000)\\
age & 0.000 & 0.000 & 0.000 & 0.001\\
 & (0.003) & (0.002) & (0.002) & (0.003)\\
marriedsingle & -0.022 & -0.027 & -0.028 & -0.023\\
 & (0.018) & (0.014) & (0.013) & (0.019)\\
\midrule
Num.Obs. & 1669 & 2229 & 2229 & 2229\\
Num.Imp. &  &  &  & 10\\
R2 & 0.208 & 0.146 & 0.223 & 0.233\\
R2 Adj. & 0.206 & 0.144 & 0.221 & 0.231\\
AIC & 1179.9 & 1093.8 & 957.6 & \\
BIC & 1223.2 & 1139.5 & 1003.2 & \\
Log.Lik. & -581.936 & -538.912 & -470.776 & \\
F & 72.917 & 63.461 & 106.192 & \\
\bottomrule
\end{tabular}
\end{table}

\exercise
We talked about this in class, but I haven't come up with a concrete idea for my project yet. I looked into the gapmindr package, but unfortunately it's not available under the current version of R. I want to do something international development related, ideally within the region of Latin America. The World Bank and UN Development Programme have great amounts of data, it just might be a bit tedious cleaning and converting from excel files to tibbles. I definitely want to include visualizations in my project this semester, as development reports typically include a lot of visualizations. But, for specific topics, I'm still undecided.

\end{document}
